Make data files

### 2021-11-27 make_data.Rmd
Create graphlist

```{r setup, include=FALSE}
library(tibble)
library(dplyr)
library(tidyr)
library(stringi)
library(lubridate)
library(intervals)
library(igraph)
requireNamespace("here")

@@

The exact way in which the `gl200210.Rdata` was created is not 100% recovered thus we create the data files based on `wgl.Rdata`. Thus some of the code below is not executed until the full pipeline is re-checked.

# Trade data

## From source files

> Code in this section is not run.

Load original data files to separate environments (I'm not sure what object names are and where):

```{r load-data, eval = FALSE}
adata <- new.env()
load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=adata)
ls(adata)

tdata <- new.env()
load(here::here("data-src", "data", "NBA_Transactions.RData"), envir=tdata)
ls(tdata)

mdata <- new.env()
load(here::here("data", "NBA_Meta.RData"), envir=mdata)
ls(mdata)

# Add rowid for back checking
main.df <- tdata$main.df %>%
  as_tibble() %>%
  tibble::rowid_to_column(".rowid")
```

Load standings data for team and division names:

```{r load-standings, eval = FALSE}
standings <- readRDS(here::here("data", "standings.rds")) %>%
  mutate(
    team_name = map_chr(strsplit(full_team_name, " "), tail, 1),
    Team = recode(team_name, "SuperSonics"="Sonics") # For joining with `standings`
  )
```

Create edgelist

```{r tradedf, eval=FALSE}
trade.df <- main.df %>%
  as_tibble() %>%
  select(.rowid, Date, Team, Notes) %>%
  mutate( 
    # Trade has "trade with" in `Notes` field...
    is_trade = grepl("trade with ", Notes),
    # ... but does not have the following phrases
    has_cancelled = grepl("cancelled ", Notes),
    has_rescinded = grepl("rescinded ", Notes),
    has_nullified = grepl("nullified ", Notes),
    has_voided = grepl("voided ", Notes)
  ) %>%
  filter(is_trade) %>%
  filter_at(vars(starts_with("has_")), all_vars(!.)) %>%
  select(-starts_with("has_"), -is_trade) %>%
  mutate( 
    # Remove stuff in parentheses from `Notes`
    notes = gsub("\\s*\\([^\\)]+\\)","", Notes),
    # Extract unique words starting with capital letter or digit (prospective team names)
    partner = map(stri_extract_all_regex(
      notes,
      paste0('\\b', paste0(unique(standings$Team), collapse="|"), '\\b'),
    ), unique),
    # Count extracted names
    num_trade_partners = map_int(partner, length)
  ) %>%
  transmute(
    team1 = Team,
    team2 = partner,
    date = Date,
    id = .rowid,
  ) %>%
  unnest("team2")
```

## Load seasonal weighted igraph objects

List of seasonal weighted igraph objects should be created from the items above. At this moment we just load it and create edge and vertex dataframes from it.

```{r load-wgl}
load(here::here("data", "wgl.Rdata"))
```

>>>>

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringi)
library(lubridate)
library(intervals)
library(igraph)
requireNamespace("here")

@@

# Trades data

Load trades data from Zachary Richardson. Group trades into seasons using April 1 of each year.

```{r}
e <- new.env()
load(here::here("data-src", "data", "NBA_AnalysisData.RData"), envir=e)

trades <- e$trade.df_long |>
  select(-season) |> # This one is wrong
  mutate(
    season_from = case_when(
      month(Date) >= 4 ~ as.numeric(year(Date)),
      month(Date) <= 3 ~ as.numeric(year(Date)) - 1,
      TRUE ~ as.numeric(NA)
    ),
    season_to = season_from + 1,
    season = paste0(season_from, "-", season_to)
  )
```

# Seasonal networks

Create a list of igraph objects with edge attribute `weight` counting the trades:

```{r graphlist}
graphlist <- lapply(
  sort(unique(trades$season)),
  function(s) {
    g <- graph_from_data_frame( 
      edb %>%
        filter(season == s) %>%
        select(P1, P2), 
      directed=FALSE)
    E(g)$weight <- count_multiple(g)
    g <- simplify(g)
    g$season <- s
    g
  }
)
```

Since that was built from edges, non-trading teams are inadvertently excluded. Here, we're adding them back in.

```{r graphlist-add-nodes}
graphlist[[3]] <- add.vertices(graphlist[[3]], 1, name="Spurs", label="Spurs", div="Central")

graphlist[[4]] <- add.vertices(graphlist[[4]], 1, name="Celtics", label="Celtics", div="Atlantic")
graphlist[[4]] <- add.vertices(graphlist[[4]], 1, name="Warriors", label="Warriors", div="Pacific")

graphlist[[5]] <- add.vertices(graphlist[[5]], 1, name="76ers", label="76ers", div="Atlantic")

graphlist[[9]] <- add.vertices(graphlist[[9]], 1, name="Nets", label="Nets", div="Atlantic")
graphlist[[9]] <- add.vertices(graphlist[[9]], 1, name="Suns", label="Suns", div="Pacific")

graphlist[[10]] <- add.vertices(graphlist[[10]], 1, name="Nuggets", label="Nuggets", div="Midwest")
graphlist[[10]] <- add.vertices(graphlist[[10]], 1, name="Knicks", label="Knicks", div="Atlantic")

graphlist[[11]] <- add.vertices(graphlist[[11]], 1, name="Suns", label="Suns", div="Pacific")
graphlist[[11]] <- add.vertices(graphlist[[11]], 1, name="Warriors", label="Warriors", div="Pacific")

graphlist[[12]] <- add.vertices(graphlist[[12]], 1, name="Mavericks", label="Mavericks", div="Midwest")

graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Lakers", label="Lakers", div="Pacific")
graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Wizards", label="Bullets", div="Atlantic")
graphlist[[14]] <- add.vertices(graphlist[[14]], 1, name="Jazz", label="Jazz", div="Midwest")

graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Pacers", label="Pacers", div="Central")
graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Celtics", label="Celtics", div="Atlantic")
graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Knicks", label="Knicks", div="Atlantic")
graphlist[[15]] <- add.vertices(graphlist[[15]], 1, name="Cavaliers", label="Cavaliers", div="Central")

graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="76ers", label="76ers", div="Atlantic")
graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Pacers", label="Pacers", div="Central")
graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Nets", label="Nets", div="Atlantic")
graphlist[[16]] <- add.vertices(graphlist[[16]], 1, name="Magic", label="Magic", div="Atlantic")

graphlist[[18]] <- add.vertices(graphlist[[18]], 1, name="Heat", label="Heat", div="Atlantic")
graphlist[[18]] <- add.vertices(graphlist[[18]], 1, name="Celtics", label="Celtics", div="Atlantic")

graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Kings", label="Kings", div="Pacific")
graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Nets", label="Nets", div="Atlantic")
graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Bulls", label="Bulls", div="Central")
graphlist[[19]] <- add.vertices(graphlist[[19]], 1, name="Cavaliers", label="Cavaliers", div="Central")

graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Pacers", label="Pacers", div="Central")
graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Rockets", label="Rockets", div="Midwest")
graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Jazz", label="Jazz", div="Midwest")
graphlist[[20]] <- add.vertices(graphlist[[20]], 1, name="Mavericks", label="Mavericks", div="Midwest")

graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="76ers", label="76ers", div="Atlantic")
graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Bulls", label="Bulls", div="Central")
graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Cavaliers", label="Cavaliers", div="Central")
graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Spurs", label="Spurs", div="Midwest")
graphlist[[21]] <- add.vertices(graphlist[[21]], 1, name="Clippers", label="Clippers", div="Pacific")

graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Lakers", label="Lakers", div="Pacific")
graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Jazz", label="Jazz", div="Midwest")
graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Wizards", label="Wizards", div="Atlantic")
graphlist[[22]] <- add.vertices(graphlist[[22]], 1, name="Spurs", label="Spurs", div="Midwest")

graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Heat", label="Heat", div="Atlantic")
graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Pacers", label="Pacers", div="Central")
graphlist[[23]] <- add.vertices(graphlist[[23]], 1, name="Clippers", label="Clippers", div="Pacific")

graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Heat", label="Heat", div="Atlantic")
graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Jazz", label="Jazz", div="Midwest")
graphlist[[24]] <- add.vertices(graphlist[[24]], 1, name="Knicks", label="Knicks", div="Atlantic")

graphlist[[25]] <- add.vertices(graphlist[[25]], 1, name="Timberwolves", label="Timberwolves", div="Midwest")
graphlist[[25]] <- add.vertices(graphlist[[25]], 1, name="Nets", label="Nets", div="Atlantic")

graphlist[[26]] <- add.vertices(graphlist[[26]], 1, name="Jazz", label="Jazz", div="Midwest")

graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Heat", label="Heat", div="Atlantic")
graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Pacers", label="Pacers", div="Central")
graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Timberwolves", label="Timberwolves", div="Midwest")
graphlist[[27]] <- add.vertices(graphlist[[27]], 1, name="Mavericks", label="Mavericks", div="Midwest")

graphlist[[28]] <- add.vertices(graphlist[[28]], 1, name="Heat", label="Heat", div="Atlantic")
graphlist[[28]] <- add.vertices(graphlist[[28]], 1, name="Lakers", label="Lakers", div="Pacific")

graphlist[[29]] <- add.vertices(graphlist[[29]], 1, name="Timberwolves", label="Timberwolves", div="Northwest")

graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Mavericks", label="Mavericks", div="Southwest")
graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Spurs", label="Spurs", div="Southwest")
graphlist[[30]] <- add.vertices(graphlist[[30]], 1, name="Warriors", label="Warriors", div="Pacific")

graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Heat", label="Heat", div="Southeast")
graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Clippers", label="Clippers", div="Pacific")
graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Kings", label="Kings", div="Pacific")
graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Wizards", label="Wizards", div="Southeast")
graphlist[[31]] <- add.vertices(graphlist[[31]], 1, name="Knicks", label="Knicks", div="Atlantic")

graphlist[[32]] <- add.vertices(graphlist[[32]], 1, name="Bucks", label="Bucks", div="Central")

graphlist[[33]] <- add.vertices(graphlist[[33]], 1, name="Hawks", label="Hawks", div="Southeast")

graphlist[[34]] <- add.vertices(graphlist[[34]], 1, name="Pacers", label="Pacers", div="Central")

graphlist[[35]] <- add.vertices(graphlist[[35]], 1, name="Pistons", label="Pistons", div="Central")

graphlist[[36]] <- add.vertices(graphlist[[36]], 1, name="Pistons", label="Pistons", div="Central")
graphlist[[36]] <- add.vertices(graphlist[[36]], 1, name="Suns", label="Suns", div="Pacific")

graphlist[[37]] <- add.vertices(graphlist[[37]], 1, name="Spurs", label="Spurs", div="Southwest")

graphlist[[38]] <- add.vertices(graphlist[[38]], 1, name="Magic", label="Magic", div="Southeast")

graphlist[[39]] <- add.vertices(graphlist[[39]], 1, name="Warriors", label="Warriors", div="Pacific")

graphlist[[41]] <- add.vertices(graphlist[[41]], 1, name="Timberwolves", label="Timberwolves", div="Northwest")

graphlist[[42]] <- add.vertices(graphlist[[42]], 1, name="Spurs", label="Spurs", div="Southwest")

graphlist[[43]] <- add.vertices(graphlist[[43]], 1, name="Warriors", label="Warriors", div="Pacific")

graphlist[[44]] <- add.vertices(graphlist[[44]], 1, name="Raptors", label="Raptors", div="Atlantic")
```

Save

```{r graphlist-save}
saveRDS(graphlist, file = "wgl.rds")
```

# Edge and node data frames
